---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gsm.studykri

<!-- badges: start -->
<!-- badges: end -->

## Overview

{gsm.studykri} provides a new approach for calculating KRI lower and upper limits on study-level for quality monitoring in clinical trials. The method uses bootstrapping to calculate confidence intervals for a given study over-time. The confidence intervals can then be used to compare the study-level KRI against a fixed expectation or against the confidence intervals and KRI values over time of one or more reference studies. The bootstrapping method resamples a new set of sites with replacement from the original study data set.

## Installation

```{r install, eval=FALSE}
install.packages("pak")
pak::pak("Gilead-BioStats/clindata")
pak::pak("Gilead-BioStats/gsm.core")
pak::pak("Gilead-BioStats/gsm.mapping")
pak::pak("Gilead-BioStats/gsm.kri")
pak::pak("Gilead-BioStats/gsm.reporting")
pak::pak("IMPALA-Consortium/gsm.studykri")
```

## Example: Days on Study

```{r example}
library(dplyr)
library(gsm.core)
library(gsm.kri)
library(gsm.studykri)

# Load raw data from clindata package
lRaw <- list(
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_PD = clindata::ctms_protdev,
  Raw_DATAENT = clindata::edc_data_pages,
  Raw_QUERY = clindata::edc_queries,
  Raw_AE = clindata::rawplus_ae,
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_Randomization = clindata::rawplus_ixrsrand,
  Raw_LB = clindata::rawplus_lb,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp,
  Raw_STUDCOMP = clindata::rawplus_studcomp,
  Raw_VISIT = clindata::rawplus_visdt
)

# Simulate portfolio with 4 studies, AA-4 oversampled for low AE counts
lPortfolio <- SimulatePortfolio(
  lRaw = lRaw,
  nStudies = 4,
  dfConfig = tibble(
    studyid = c("AA-1", "AA-2", "AA-3", "AA-4"),
    nSubjects = c(500, 750, 150, 200),
    strOversamplDomain = rep("Raw_AE", 4),
    vOversamplQuantileRange_min = c(0, 0, 0, 0),
    vOversamplQuantileRange_max = c(1, 1, 1, 0.75)
  )
)

# Calculate site-level counts by month using days on study as denominator
# Pass subjects as denominator with start and end dates to calculate person-days
dfInputDays <- Input_CountSiteByMonth(
  dfSubjects = lPortfolio$Raw_SUBJ,
  dfNumerator = lPortfolio$Raw_AE,
  dfDenominator = lPortfolio$Raw_SUBJ,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "firstparticipantdate",
  strDenominatorEndDateCol = "lastparticipantdate"
) %>% tibble()

# Aggregate to study-level cumulative KRI by month
dfTransformed <- Transform_CumCount(
  dfInputDays,
  nMinDenominator = 25,
  vBy = c("StudyID")
) %>% tibble()

# Generate bootstrap resamples at site-level (1000 iterations)
df_Analyzed_Bootstrap_Site <- dfInputDays %>%
  Analyze_StudyKRI(nBootstrapReps = 1000) %>%
  tibble()

# Aggregate each bootstrap iteration to study-level
df_Analyzed_Bootstrap_Study <- df_Analyzed_Bootstrap_Site %>%
  Transform_CumCount(
    nMinDenominator = 25,
    vBy = c("StudyID", "BootstrapRep")
  ) %>% tibble()

# Calculate 95% confidence intervals from bootstrap distribution
df_Bounds <- Analyze_StudyKRI_PredictBounds(
  df_Analyzed_Bootstrap_Study,
  vBy = c("StudyID"),
  nConfLevel = 0.95
) %>% tibble()

# Plot KRI over time with confidence intervals for study AA-4
Visualize_StudyKRI(
  dfStudyKRI = dfTransformed,
  dfBounds = df_Bounds,
  strStudyID = "AA-4"
)
```

## AI Disclaimer

This package and documentation were developed with assistance from AI tools, including Cursor and Claude Sonnet 4.5. All AI-generated content has been reviewed.

## License

MIT © 2025 IMPALA Consortium

