---
title: "Cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookbook}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r cook1 , include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r cook2, eval=FALSE}
install.packages("pak")
pak::pak("Gilead-BioStats/clindata")
pak::pak("Gilead-BioStats/gsm.core")
pak::pak("Gilead-BioStats/gsm.mapping")
pak::pak("Gilead-BioStats/gsm.kri")
pak::pak("Gilead-BioStats/gsm.reporting")
pak::pak("IMPALA-Consortium/gsm.simaerep")
```

## Load

```{r cook3}
suppressPackageStartupMessages(library(dplyr))
library(gsm.core)
library(gsm.mapping)
library(gsm.kri)
library(gsm.reporting)
library(gsm.studykri)
```

## Introduction

{gsm.studykir} presents an new approach for calculating KRI lower and upper limits on study-level for quality monitoring in clincial trials. The method uses bootstrapping to calculate
confidence intervals gor a given study over-time. The confidence intervals can then be used to compare the study-level KRI against a fixed expectation or against the confidence intervalls
and KRI values over time of one or more reference studies. The bootstrapping method resamples a new set of sites with replacement from the original study data set.


## Simulate Data Set

We use the clindata package which includes data from one clinical study to simulate new study data. We create a portfolio
with study AA-4 oversampling patients with low AE counts.

```{r}
lRaw <- list(
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_PD = clindata::ctms_protdev,
  Raw_DATAENT = clindata::edc_data_pages,
  Raw_QUERY = clindata::edc_queries,
  Raw_AE = clindata::rawplus_ae,
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_Randomization = clindata::rawplus_ixrsrand,
  Raw_LB = clindata::rawplus_lb,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp,
  Raw_STUDCOMP = clindata::rawplus_studcomp,
  Raw_VISIT = clindata::rawplus_visdt
)

# to ensure proper resampling we need to make sure that lRaw$Raw_SUBJ 
# can be used as a proper lookup table for all subject identifier aliases

lRaw$Raw_SUBJ$subjectname <- lRaw$Raw_SUBJ$subject_nsv
lRaw$Raw_SUBJ$subjectenrollmentnumber <- lRaw$Raw_SUBJ$subjectid

lPortfolio <- SimulatePortfolio(
  lRaw = lRaw,
  nStudies = 4,
  vSubjectIDs = c("subjid", "subjectenrollmentnumber", "subject_nsv", "subjectname", "subjectid"),
  dfConfig = tibble(
    studyid = c("AA-1", "AA-2", "AA-3", "AA-4"),
    nSubjects = c(500, 750, 150, 200),
    strOversamplDomain = rep("Raw_AE", 4),
    vOversamplQuantileRange_min = c(0, 0, 0, 0),
    vOversamplQuantileRange_max = c(1, 1, 1, 0.75)
  )
)
```

## Calculate KRI

We will calculate the KRI "AE per visit". For this we need to create three dataframes:

- dfSubjects: Subject-level data that links subjects and sites from the demographics domain
- dfNumerator: Subject-level data that contains the events to be counted (AEs in this case)
- dfDenominator: Subject-level data that contains the denominator events (visits in this case

that we transform to return the site-level count of numerator and denominator per month

```{r}
dfInput <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = lPortfolio$Raw_SUBJ,
  dfNumerator = lPortfolio$Raw_AE,
  dfDenominator = lPortfolio$Raw_VISIT,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt",
)

dfInput
```

Alternatively we can also use days on study as a denominator by passing subjects as denominator and an end date.

```{r}
dfInputDays <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = lPortfolio$Raw_SUBJ,
  dfNumerator = lPortfolio$Raw_AE,
  dfDenominator = lPortfolio$Raw_SUBJ,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "firstparticipantdate",
  strDenominatorEndDateCol = "lastparticipantdate"
)

dfInputDays
```

Next we transform the data to return the cumulated event counts on study-level by study month. We define the first study month to be the month in which
we have reached the first 25 Denominator counts across all sites to account for data artifacts.

```{r}
dfTransformed <- gsm.studykri::Transform_CumCount(
  dfInput,
  nMinDenominator = 25,
  vBy = c("StudyID")
)

dfTransformed
```

Analyze_StudyKRI_PredictBounds calculates upper and lower confidence intervalls of the study metrics KRI using a bootstrap technique.

```{r}
dfBounds <- dfInput %>%
  gsm.studykri::Analyze_StudyKRI_PredictBounds()

dfBounds
```

We can plot the preliminary results.

```{r}
gsm.studykri::Visualize_StudyKRI(
  dfStudyKRI = dfTransformed ,
  dfBounds = dfBounds,
  strStudyID = "AA-1"
)
```


We then calculate a reference confidence interval based on a set of reference studies.

```{r}

dfStudyRef <- tibble(
  study = "AA-1",
  studyref = c("AA-2", "AA-3", "AA-4")
)

dfBoundsRef <- gsm.studykri::Analyze_StudyKRI_PredictBoundsRef(dfInput, dfStudyRef)

dfBoundsRef
```

And add the reference confidence interval to out plot

```{r}
gsm.studykri::Visualize_StudyKRI(
  dfStudyKRI = dfTransformed ,
  dfBounds = dfBounds,
  dfBoundsRef = dfBoundsRef,
  strStudyID = "AA-1"
)
```


We can also plot the members of the reference portfolio.

```{r}
p1 <- gsm.studykri::Visualize_StudyKRI(
  dfStudyKRI = dfTransformed ,
  dfBounds = dfBounds,
  dfBoundsRef = dfBoundsRef,
  strStudyID = "AA-2"
)

p2 <- gsm.studykri::Visualize_StudyKRI(
  dfStudyKRI = dfTransformed ,
  dfBounds = dfBounds,
  dfBoundsRef = dfBoundsRef,
  strStudyID = "AA-3"
)

p3 <- gsm.studykri::Visualize_StudyKRI(
  dfStudyKRI = dfTransformed ,
  dfBounds = dfBounds,
  dfBoundsRef = dfBoundsRef,
  strStudyID = "AA-4"
)

library(patchwork)

p1 + p2 + p3 + patchwork::plot_layout(guides = "collect")
```



## Calculate KRI using yaml workflows

We can also use the {gsm}-style yaml workflow. Here we start with a portfolio with 6 studies of
 which 2 are target studies for whcih reports will be built, while the others serve as reference.

```{r cook4}
# Generate multi-study portfolio using SimulatePortfolio
lRaw_original <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt,
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_PD = clindata::ctms_protdev,
  Raw_DATAENT = clindata::edc_data_pages,
  Raw_DATACHG = clindata::edc_data_points,
  Raw_QUERY = clindata::edc_queries,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_Randomization = clindata::rawplus_ixrsrand,
  Raw_LB = clindata::rawplus_lb,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp,
  Raw_STUDCOMP = clindata::rawplus_studcomp,
  Raw_IE = clindata::rawplus_ie
)

lRaw_original$Raw_SUBJ$subjectname <- lRaw_original$Raw_SUBJ$subject_nsv
lRaw_original$Raw_SUBJ$subjectenrollmentnumber <- lRaw_original$Raw_SUBJ$subjid

# Create a portfolio with 3 studies
lRaw <- SimulatePortfolio(
  lRaw = lRaw_original,
  nStudies = 3,
  seed = 123,
  vSubjectIDs = c("subjid", "subjectenrollmentnumber", "subject_nsv", "subjectname", "subjectid"),
  dfConfig = tibble(
    studyid = c("AA-1", "AA-2", "AA-3", "AA-4", "AA-5", "AA-6"),
    nSubjects = c(500, 750, 150, 200, 250, 300),
    strOversamplDomain = rep("Raw_AE", 6),
    vOversamplQuantileRange_min = c(0, 0, 0, 0, 0, 0),
    vOversamplQuantileRange_max = c(1, 0.75, 1, 1, 0.95, 0.9)
  )
)

lRaw$Raw_StudyRef <- tibble(
  studyid = c(rep("AA-1", 3), rep("AA-2", 4)),
  studyrefid = c("AA-3", "AA-4", "AA-5", "AA-3", "AA-4", "AA-5", "AA-6")
)

# Run mapping workflows
mapping_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/1_mappings", package = "gsm.studykri"),
  strPackage = NULL
)

lIngest <- gsm.mapping::Ingest(lRaw, gsm.mapping::CombineSpecs(mapping_wf))


lMapped <- gsm.core::RunWorkflows(lWorkflows = mapping_wf, lData = lIngest)


# Run KRI workflow
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

# Combine lMapped with metrics workflows for comprehensive result
lAnalyzed <- gsm.core::RunWorkflows(lWorkflows = metrics_wf, lData = lMapped)

```

## Generate KRI report using scripts

and subsequently generate a KRI report using {gsm.reporting}.

We assemble some data needed for reporting.

```{r cook5}

dfInput <- gsm.studykri::BindResults(lAnalyzed, "Analysis_Input")

dfTransformed <- gsm.studykri::BindResults(lAnalyzed, "Analysis_Transformed")

dfMetrics <- gsm.reporting::MakeMetric(metrics_wf)

dfGroups <- dplyr::bind_rows(
  lMapped$Mapped_STUDY,
  lMapped$Mapped_SITE,
  lMapped$Country
)

dfStudyRef <- lMapped$Mapped_StudyRef

dfStudyRef
```

In order to make the bootstrapping more efficient we can combine all KRI that use the same denominator.

```{r}

lJoined <- gsm.studykri::JoinKRIByDenominator(dfInput)

str(lJoined)

```

Using the study references we can calculate the bootstrapped confidence intervals for all KRI at once.

```{r}
Bounds_Wide <- purrr::map(lJoined, ~ Analyze_StudyKRI_PredictBounds(., dfStudyRef))
BoundsRef_Wide <- purrr::map(lJoined, ~ Analyze_StudyKRI_PredictBoundsRef(., dfStudyRef))

dfBounds <- Transform_Long(Bounds_Wide)
dfBoundsRef <- Transform_Long(BoundsRef_Wide)
```

```{r}

lCharts <- gsm.studykri::MakeCharts_StudyKRI(
    dfResults = dfTransformed, 
    dfBounds = dfBounds,
    dfBoundsRef = dfBoundsRef,
    dfMetrics = dfMetrics
  )

gsm.studykri::Report_StudyKRI(
  lCharts = lCharts,
  dfResults = dfTransformed,
  dfGroups = dfGroups,
  dfMetrics = dfMetrics,
  strOutputFile = "report_studykri.html",
  strInputPath = system.file("report", "Report_KRI.Rmd", package = "gsm.studykri")
)
```

## Generate KRI report using yaml workflows

```{r cook6}
reporting_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_reporting", package = "gsm.studykri")
)

# Pass both lMapped and lAnalyzed, plus workflows
lReporting <- gsm.core::RunWorkflows(
  lWorkflows = reporting_wf,
  lData = c(
    lMapped,
    list(
      lAnalyzed = lAnalyzed,
      lWorkflows = metrics_wf
    )
  )
)

module_wf_gsm <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/4_modules", package = "gsm.studykri"),
  strPackage = NULL
)

# we cannot set a dynamic link to the report path in the yaml files
report_path <- system.file("report", "Report_KRI.Rmd", package = "gsm.studykri")
n_steps <- length(module_wf_gsm$StudyKRI$steps)

module_wf_gsm$StudyKRI$steps[[n_steps]]$params$strInputPath <- report_path

lModule <- gsm.core::RunWorkflows(module_wf_gsm, lReporting)
```
