---
title: "In-Database Compatibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{In-Database Compatibility}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r cook1 , include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r cook2, eval=FALSE}
install.packages("pak")
pak::pak("Gilead-BioStats/clindata")
pak::pak("Gilead-BioStats/gsm.core")
pak::pak("Gilead-BioStats/gsm.mapping")
pak::pak("Gilead-BioStats/gsm.kri")
pak::pak("Gilead-BioStats/gsm.reporting")
pak::pak("IMPALA-Consortium/gsm.simaerep")
```

## Load

```{r cook3}
suppressPackageStartupMessages(library(dplyr))
library(gsm.core)
library(gsm.mapping)
library(gsm.kri)
library(gsm.reporting)
library(gsm.studykri)
```

## Introduction

{gsm.studykir} functions can be executed on a database back-end instead of relying on in-memory calculations. 

The package supports various database backends including DuckDB, PostgreSQL, and Snowflake. For backends that use large integer values for random number generation (such as Snowflake), use the `vDbIntRandomRange` parameter in bootstrap functions to ensure proper normalization of random values.

## Database Backend Compatibility

### Random Number Generation

Most SQL databases translate R's `runif()` to a `RANDOM()` function that returns decimal values between 0 and 1. However, some databases (notably Snowflake) return large integers instead. For these backends, you must specify the integer range using the `vDbIntRandomRange` parameter.

**Common Random Number Ranges:**

- **Snowflake** (signed 64-bit): `c(-9223372036854775808, 9223372036854775807)` (range: −2^63 to 2^63−1)
- **Standard databases** (DuckDB, PostgreSQL, etc.): `NULL` (default, returns 0-1 decimals)

When `vDbIntRandomRange` is specified, the package automatically normalizes the integer values to the 0-1 range using the formula:

```r
normalized_value = (random_value - min_value) / (max_value - min_value)
```

**How to determine your database's random range:**

If you're unsure whether your database backend needs normalization, run this query:

```sql
SELECT RANDOM() FROM your_table LIMIT 10;
```

If the values are between 0 and 1, use the default (`vDbIntRandomRange = NULL`). If they're large integers, specify the appropriate range.


## Simulate Data Set and Save to DuckDb

We use the clindata package which includes data from one clinical study to simulate new study data. We create a portfolio
with study AA-4 oversampling patients with low AE counts.

```{r}
lRaw <- list(
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_PD = clindata::ctms_protdev,
  Raw_DATAENT = clindata::edc_data_pages,
  Raw_DATACHG = clindata::edc_data_points,
  Raw_QUERY = clindata::edc_queries,
  Raw_AE = clindata::rawplus_ae,
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_Randomization = clindata::rawplus_ixrsrand,
  Raw_LB = clindata::rawplus_lb,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp,
  Raw_STUDCOMP = clindata::rawplus_studcomp,
  Raw_IE = clindata::rawplus_ie,
  Raw_VISIT = clindata::rawplus_visdt %>%
    mutate(visit_dt = lubridate::ymd(visit_dt))
)

# Add required subject identifier columns for resampling
lRaw$Raw_SUBJ$subjectname <- lRaw$Raw_SUBJ$subject_nsv
lRaw$Raw_SUBJ$subjectenrollmentnumber <- lRaw$Raw_SUBJ$subjid

lPortfolio <- SimulatePortfolio(
  lRaw = lRaw,
  nStudies = 4,
  vSubjectIDs = c("subjid", "subjectenrollmentnumber", "subject_nsv", "subjectname", "subjectid"),
  dfConfig = tibble(
    studyid = c("AA-1", "AA-2", "AA-3", "AA-4"),
    nSubjects = c(500, 750, 150, 200),
    strOversampleDomain = rep("Raw_AE", 4),
    vOversamplQuantileRange_min = c(0, 0, 0, 0),
    vOversamplQuantileRange_max = c(1, 1, 1, 0.75)
  )
)

con <- DBI::dbConnect(duckdb::duckdb(), ":memory:")

purrr::walk2(names(lPortfolio), lPortfolio, ~ DBI::dbWriteTable(con, .x, .y))

dbPortfolio <- purrr::map(names(lPortfolio), ~ tbl(con, .x))
names(dbPortfolio) <- names(lPortfolio)
```

Here we create our data set.

```{r}
tblInput1 <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = dbPortfolio$Raw_SUBJ,
  dfNumerator = dbPortfolio$Raw_AE,
  dfDenominator = dbPortfolio$Raw_VISIT,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt",
  strDenominatorType = "Visit"
)

tblInput2 <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = dbPortfolio$Raw_SUBJ,
  dfNumerator = dbPortfolio$Raw_AE %>%
    filter(aeser == "Y"),
  dfDenominator = dbPortfolio$Raw_VISIT,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt",
  strDenominatorType = "Visit"
)

tblInput <- dplyr::union_all(
  tblInput1 %>%
    mutate(MetricID = "kri0001"),
  tblInput2 %>%
    mutate(MetricID = "kri0002"),
)

# Load metric metadata from workflow files
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = c("kri0001.yaml", "kri0002.yaml"),
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

dfMetrics <- gsm.reporting::MakeMetric(metrics_wf)

lJoined <- JoinKRIByDenominator(tblInput, dfMetrics)

lJoined
```


Next we prepare for the actual bootstrap simulation. For this we need to perform multiple cross joins that require the creation of
temporary tables in the backend. {gsm.studykri} will attempt to create them using the connection attached to the lazy tables.
However if write priveleges are not sufficient we can also create the tables manually and pass them to the relevant functions.

```{r}
dfStudyRef <- tibble(
  StudyID = "AA-1",
  StudyRef = c("AA-2", "AA-3", "AA-4")
)

dfRep <- tibble(
  BootstrapRep = seq(1, 1000)
)

dfMonths <- tibble(
  YYYY = seq(2003, 2020)
) %>%
  cross_join(
    tibble(
      MM = seq(1, 12)
    )
  ) %>%
  mutate(
    MonthYYYYMM = YYYY * 100 + MM
  ) %>%
  select(MonthYYYYMM)

DBI::dbWriteTable(con, "StudyRef", dfStudyRef, overwrite = TRUE)
DBI::dbWriteTable(con, "Rep", dfRep, overwrite = TRUE)
DBI::dbWriteTable(con, "Months", dfMonths, overwrite = TRUE)


tblStudyRef <- tbl(con, "StudyRef")
tblRep <- tbl(con, "Rep")
tblMonths <- tbl(con, "Months")


Bounds_Wide <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBounds(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths
  )
)

BoundsRef_Wide <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBoundsRef(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths
  )
)

tblBounds <- Transform_Long(Bounds_Wide)

tblBoundsRef <- Transform_Long(BoundsRef_Wide)

tblTransformed <- union_all(
  tblInput1 %>%
    Transform_CumCount(vBy = "StudyID", tblMonthSequence = tblMonths) %>%
    mutate(MetricID = "kri0001"),
  tblInput2 %>%
    Transform_CumCount(vBy = "StudyID", tblMonthSequence = tblMonths) %>%
    mutate(MetricID = "kri0002"),
)


tblBounds

tblBoundsRef

tblTransformed
```

## Using Snowflake or Other Integer Random Backends

If you're using Snowflake or another database that returns large integers for `RANDOM()`, add the `vDbIntRandomRange` parameter to your bootstrap function calls:

```{r eval=FALSE}
# For Snowflake: specify signed 64-bit integer range
Bounds_Wide_Snowflake <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBounds(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths,
    vDbIntRandomRange = c(-9223372036854775808, 9223372036854775807)
  )
)

BoundsRef_Wide_Snowflake <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBoundsRef(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths,
    vDbIntRandomRange = c(-9223372036854775808, 9223372036854775807)
  )
)
```

The `vDbIntRandomRange` parameter should be added to:

- `Analyze_StudyKRI_PredictBounds()`
- `Analyze_StudyKRI_PredictBoundsRef()`
- `Analyze_StudyKRI_PredictBoundsRefSet()`

This ensures that random values are properly normalized regardless of the database backend's random number generation method.

## Performance Optimization for Database Backends

### Overview of Collect Operations

When working with database backends, certain operations require materializing data from the database into R memory using `collect()`. Understanding these operations helps optimize performance:

| Location | Purpose | Data Size | Avoidable? | How to Avoid |
|----------|---------|-----------|------------|--------------|
| **`Analyze_StudyKRI_PredictBoundsRefSet`** | Calculate min group count | Full filtered dataset | **Yes** | Add MinGroups column to StudyRef |
| `Analyze_StudyKRI_PredictBoundsRef` | Extract study mappings | Small (< 100 rows) | No - needed for R loop | N/A |
| `Transform_CumCount` | Get min/max MonthYYYYMM | 2 numeric values | No - minimal impact | N/A |
| `JoinKRIByDenominator` | Validation checks | Varies | Yes | Use `bSkipValidation = TRUE` |
| Final reporting | Create charts | Full result set | No - required | N/A |

### Optimization Strategy 1: Pass nMinGroups via StudyRef

The most impactful optimization is avoiding the collection of the full dataset to calculate minimum group counts. By pre-calculating and including a MinGroups column in your StudyRef table, you can avoid materializing potentially millions of rows.

**How to calculate MinGroups:**

For each target study in your StudyRef mapping, calculate the minimum group count across all its reference studies:

```sql
-- Example: Calculate MinGroups for each target-reference set
WITH GroupCounts AS (
  SELECT 
    StudyID,
    COUNT(DISTINCT GroupID) AS group_count
  FROM Analysis_Input
  WHERE StudyID IN ('REF_STUDY_1', 'REF_STUDY_2', 'REF_STUDY_3')
  GROUP BY StudyID
)
SELECT 
  'TARGET_STUDY' AS studyid,
  'REF_STUDY_1' AS studyrefid,  -- repeat for each reference
  MIN(group_count) AS MinGroups
FROM GroupCounts;
```

**Then include it in your StudyRef table:**

The StudyRef.yaml workflow selects only the required `studyid` and `studyrefid` columns. To use the MinGroups optimization, you have two options:

**Option 1: Add MinGroups directly to your StudyRef data**

```{r eval=FALSE}
# Create Raw_StudyRef with MinGroups column
dfStudyRef <- tibble(
  studyid = "AA-1",
  studyrefid = c("AA-2", "AA-3", "AA-4"),
  MinGroups = 8 # Pre-calculated minimum across reference studies
)

# Write to database
DBI::dbWriteTable(con, "StudyRef", dfStudyRef, overwrite = TRUE)
tblStudyRef <- tbl(con, "StudyRef")

# Use in analysis - MinGroups will be automatically detected and used
BoundsRef_Optimized <- Analyze_StudyKRI_PredictBoundsRef(
  dfInput = tblInput,
  dfStudyRef = tblStudyRef,
  tblBootstrapReps = tblRep,
  tblMonthSequence = tblMonths
)
```

**Option 2: Use a custom mapping query**

If you need to transform the column name, create a custom StudyRef.yaml workflow:

```yaml
steps:
  - output: Mapped_StudyRef
    name: gsm.core::RunQuery
    params:
      df: Raw_StudyRef
      strQuery: "SELECT studyid AS study, studyrefid AS studyref, mingroups AS MinGroups FROM df"
```

**Option 3: Use a custom column name**

```{r eval=FALSE}
dfStudyRef_Custom <- tibble(
  studyid = "AA-1",
  studyrefid = c("AA-2", "AA-3", "AA-4"),
  min_grp_cnt = 8
)

BoundsRef_Custom <- Analyze_StudyKRI_PredictBoundsRef(
  dfInput = tblInput,
  dfStudyRef = tblStudyRef_Custom,
  strMinGroupsCol = "min_grp_cnt", # Specify custom column name
  tblBootstrapReps = tblRep,
  tblMonthSequence = tblMonths
)
```

**Performance Impact:**

- **Before**: Collects full filtered dataset → groups by StudyID and GroupID → calculates minimum
- **After**: Reads single integer value from StudyRef (no collect on main dataset)
- **Savings**: Can avoid collecting millions of rows for large studies

### Optimization Strategy 2: Skip Validation Checks

The `JoinKRIByDenominator()` function performs several validation checks that require `collect()` operations. When confident in data quality, skip these:

```{r eval=FALSE}
lJoined <- JoinKRIByDenominator(
  tblInput,
  dfMetrics,
  bSkipValidation = TRUE # Disable validation collect() operations
)
```

### Optimization Strategy 3: Pre-generate Temporary Tables

Provide pre-generated bootstrap replicates and month sequences to avoid repeated temp table creation:

```{r eval=FALSE}
# Create once, reuse many times
dfRep <- tibble(BootstrapRep = seq(1, 1000))
dfMonths <- tibble(
  YYYY = seq(2003, 2020)
) %>%
  cross_join(tibble(MM = seq(1, 12))) %>%
  mutate(MonthYYYYMM = YYYY * 100 + MM) %>%
  select(MonthYYYYMM)

DBI::dbWriteTable(con, "Rep", dfRep)
DBI::dbWriteTable(con, "Months", dfMonths)

tblRep <- tbl(con, "Rep")
tblMonths <- tbl(con, "Months")

# Pass to all functions
Bounds <- Analyze_StudyKRI_PredictBounds(
  dfInput,
  tblBootstrapReps = tblRep,
  tblMonthSequence = tblMonths
)
```

### Best Practices Summary

1. **Add MinGroups column to StudyRef** (or use `strMinGroupsCol` parameter for custom names) for the biggest performance gain
2. **Use `bSkipValidation = TRUE`** in `JoinKRIByDenominator()` when data quality is assured
3. **Pre-generate temp tables** (`tblBootstrapReps`, `tblMonthSequence`) to avoid repeated creation
4. **Ensure proper indexing** on StudyID, GroupID, and MonthYYYYMM columns in your database
5. **Collect only at the end** - keep data as lazy tables until final visualization/reporting

Finally before plotting we need to collect the results into memory.

```{r}
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

dfMetrics <- gsm.reporting::MakeMetric(metrics_wf)


lCharts <- gsm.studykri::MakeCharts_StudyKRI(
  dfResults = collect(tblTransformed),
  dfBounds = collect(tblBounds),
  dfBoundsRef = collect(tblBoundsRef),
  dfMetrics = dfMetrics
)

lCharts
```

## Using workflows

Unfortunately the data specfication checks from gsm.core will not work on lazy tables. We therefore need to skip the 
mapping workflow and ignore the warnings for the missing specifications.

We might also scrub some of the specifications from the yaml files to begin with.

```{r}
lMapped <- list(
  Mapped_AE = dbPortfolio$Raw_AE,
  Mapped_Visit = dbPortfolio$Raw_VISIT,
  Mapped_SUBJ = dbPortfolio$Raw_SUBJ %>%
    filter(enrollyn == "Y"),
  Mapped_StudyRef = tbl(con, "StudyRef")
)

metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = "kri0001.yaml",
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

lAnalyzed <- gsm.core::RunWorkflows(lWorkflows = metrics_wf, lData = lMapped)

reporting_wf <- gsm.core::MakeWorkflowList(
  strNames = c("Bounds", "BoundsRef", "Input", "Join", "Metrics", "Results"),
  strPath = system.file("workflow/3_reporting", package = "gsm.studykri")
)

lReporting <- gsm.core::RunWorkflows(
  lWorkflows = reporting_wf,
  lData = c(
    lMapped,
    list(
      lAnalyzed = lAnalyzed,
      lWorkflows = metrics_wf
    )
  )
)

lReporting
```

Next we collect the data and create the charts.

```{r}
lCharts <- gsm.studykri::MakeCharts_StudyKRI(
  dfResults = collect(lReporting$Reporting_Results),
  dfBounds = collect(lReporting$Reporting_Bounds),
  dfBoundsRef = collect(lReporting$Reporting_BoundsRef),
  dfMetrics = lReporting$Reporting_Metrics
)

lCharts
```
