---
title: "In-Database Compatibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{In-Database Compatibility}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r cook1 , include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r cook2, eval=FALSE}
install.packages("pak")
pak::pak("Gilead-BioStats/clindata")
pak::pak("Gilead-BioStats/gsm.core")
pak::pak("Gilead-BioStats/gsm.mapping")
pak::pak("Gilead-BioStats/gsm.kri")
pak::pak("Gilead-BioStats/gsm.reporting")
pak::pak("IMPALA-Consortium/gsm.simaerep")
```

## Load

```{r cook3}
suppressPackageStartupMessages(library(dplyr))
library(gsm.core)
library(gsm.mapping)
library(gsm.kri)
library(gsm.reporting)
library(gsm.studykri)
```

## Introduction

{gsm.studykir} functions can be executed on a database back-end instead of relying on in-memory calculations. 

The package supports various database backends including DuckDB, PostgreSQL, and Snowflake. For backends that use large integer values for random number generation (such as Snowflake), use the `vDbIntRandomRange` parameter in bootstrap functions to ensure proper normalization of random values.

## Database Backend Compatibility

### Random Number Generation

Most SQL databases translate R's `runif()` to a `RANDOM()` function that returns decimal values between 0 and 1. However, some databases (notably Snowflake) return large integers instead. For these backends, you must specify the integer range using the `vDbIntRandomRange` parameter.

**Common Random Number Ranges:**

- **Snowflake** (signed 64-bit): `c(-9223372036854775808, 9223372036854775807)` (range: −2^63 to 2^63−1)
- **Standard databases** (DuckDB, PostgreSQL, etc.): `NULL` (default, returns 0-1 decimals)

When `vDbIntRandomRange` is specified, the package automatically normalizes the integer values to the 0-1 range using the formula:

```r
normalized_value = (random_value - min_value) / (max_value - min_value)
```

**How to determine your database's random range:**

If you're unsure whether your database backend needs normalization, run this query:

```sql
SELECT RANDOM() FROM your_table LIMIT 10;
```

If the values are between 0 and 1, use the default (`vDbIntRandomRange = NULL`). If they're large integers, specify the appropriate range.


## Simulate Data Set and Save to DuckDb

We use the clindata package which includes data from one clinical study to simulate new study data. We create a portfolio
with study AA-4 oversampling patients with low AE counts.

```{r}
lRaw <- list(
  Raw_SITE = clindata::ctms_site,
  Raw_STUDY = clindata::ctms_study,
  Raw_PD = clindata::ctms_protdev,
  Raw_DATAENT = clindata::edc_data_pages,
  Raw_DATACHG = clindata::edc_data_points,
  Raw_QUERY = clindata::edc_queries,
  Raw_AE = clindata::rawplus_ae,
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_ENROLL = clindata::rawplus_enroll,
  Raw_Randomization = clindata::rawplus_ixrsrand,
  Raw_LB = clindata::rawplus_lb,
  Raw_SDRGCOMP = clindata::rawplus_sdrgcomp,
  Raw_STUDCOMP = clindata::rawplus_studcomp,
  Raw_IE = clindata::rawplus_ie,
  Raw_VISIT = clindata::rawplus_visdt %>%
    mutate(visit_dt = lubridate::ymd(visit_dt))
)

# Add required subject identifier columns for resampling
lRaw$Raw_SUBJ$subjectname <- lRaw$Raw_SUBJ$subject_nsv
lRaw$Raw_SUBJ$subjectenrollmentnumber <- lRaw$Raw_SUBJ$subjid

lPortfolio <- SimulatePortfolio(
  lRaw = lRaw,
  nStudies = 4,
  vSubjectIDs = c("subjid", "subjectenrollmentnumber", "subject_nsv", "subjectname", "subjectid"),
  dfConfig = tibble(
    studyid = c("AA-1", "AA-2", "AA-3", "AA-4"),
    nSubjects = c(500, 750, 150, 200),
    strOversamplDomain = rep("Raw_AE", 4),
    vOversamplQuantileRange_min = c(0, 0, 0, 0),
    vOversamplQuantileRange_max = c(1, 1, 1, 0.75)
  )
)

con <- DBI::dbConnect(duckdb::duckdb(), ":memory:")

purrr::walk2(names(lPortfolio), lPortfolio, ~ DBI::dbWriteTable(con, .x, .y))

dbPortfolio <- purrr::map(names(lPortfolio), ~ tbl(con, .x))
names(dbPortfolio) <- names(lPortfolio)
```

Here we create our data set.

```{r}
tblInput1 <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = dbPortfolio$Raw_SUBJ,
  dfNumerator = dbPortfolio$Raw_AE,
  dfDenominator = dbPortfolio$Raw_VISIT,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt",
  strDenominatorType = "Visit"
)

tblInput2 <- gsm.studykri::Input_CountSiteByMonth(
  dfSubjects = dbPortfolio$Raw_SUBJ,
  dfNumerator = dbPortfolio$Raw_AE %>%
    filter(aeser == "Y"),
  dfDenominator = dbPortfolio$Raw_VISIT,
  strStudyCol = "studyid",
  strGroupCol = "invid",
  strGroupLevel = "Site",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt",
  strDenominatorType = "Visit"
)

tblInput <- dplyr::union_all(
  tblInput1 %>%
    mutate(MetricID = "kri0001"),
  tblInput2 %>%
    mutate(MetricID = "kri0002"),
)

# Load metric metadata from workflow files
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = c("kri0001.yaml", "kri0002.yaml"),
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

dfMetrics <- gsm.reporting::MakeMetric(metrics_wf)

lJoined <- JoinKRIByDenominator(tblInput, dfMetrics)

lJoined
```


Next we prepare for the actual bootstrap simulation. For this we need to perform multiple cross joins that require the creation of
temporary tables in the backend. {gsm.studykri} will attempt to create them using the connection attached to the lazy tables.
However if write priveleges are not sufficient we can also create the tables manually and pass them to the relevant functions.

```{r}
dfStudyRef <- tibble(
  StudyID = "AA-1",
  StudyRef = c("AA-2", "AA-3", "AA-4")
)

dfRep <- tibble(
  BootstrapRep = seq(1, 1000)
)

dfMonths <- tibble(
  YYYY = seq(2003, 2020)
) %>%
  cross_join(
    tibble(
      MM = seq(1, 12)
    )
  ) %>%
  mutate(
    MonthYYYYMM = YYYY * 100 + MM
  ) %>%
  select(MonthYYYYMM)

DBI::dbWriteTable(con, "StudyRef", dfStudyRef, overwrite = TRUE)
DBI::dbWriteTable(con, "Rep", dfRep, overwrite = TRUE)
DBI::dbWriteTable(con, "Months", dfMonths, overwrite = TRUE)


tblStudyRef <- tbl(con, "StudyRef")
tblRep <- tbl(con, "Rep")
tblMonths <- tbl(con, "Months")


Bounds_Wide <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBounds(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths
  )
)

BoundsRef_Wide <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBoundsRef(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths
  )
)

tblBounds <- Transform_Long(Bounds_Wide)

tblBoundsRef <- Transform_Long(BoundsRef_Wide)

tblTransformed <- union_all(
  tblInput1 %>%
    Transform_CumCount(vBy = "StudyID", tblMonthSequence = tblMonths) %>%
    mutate(MetricID = "kri0001"),
  tblInput2 %>%
    Transform_CumCount(vBy = "StudyID", tblMonthSequence = tblMonths) %>%
    mutate(MetricID = "kri0002"),
)


tblBounds

tblBoundsRef

tblTransformed
```

## Using Snowflake or Other Integer Random Backends

If you're using Snowflake or another database that returns large integers for `RANDOM()`, add the `vDbIntRandomRange` parameter to your bootstrap function calls:

```{r eval=FALSE}
# For Snowflake: specify signed 64-bit integer range
Bounds_Wide_Snowflake <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBounds(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths,
    vDbIntRandomRange = c(-9223372036854775808, 9223372036854775807)
  )
)

BoundsRef_Wide_Snowflake <- purrr::map(
  lJoined, ~ Analyze_StudyKRI_PredictBoundsRef(
    .,
    dfStudyRef = tblStudyRef,
    tblBootstrapReps = tblRep,
    tblMonthSequence = tblMonths,
    vDbIntRandomRange = c(-9223372036854775808, 9223372036854775807)
  )
)
```

The `vDbIntRandomRange` parameter should be added to:

- `Analyze_StudyKRI_PredictBounds()`
- `Analyze_StudyKRI_PredictBoundsRef()`
- `Analyze_StudyKRI_PredictBoundsRefSet()`

This ensures that random values are properly normalized regardless of the database backend's random number generation method.

Finally before plotting we need to collect the results into memory.

```{r}
metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = NULL,
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

dfMetrics <- gsm.reporting::MakeMetric(metrics_wf)


lCharts <- gsm.studykri::MakeCharts_StudyKRI(
  dfResults = collect(tblTransformed),
  dfBounds = collect(tblBounds),
  dfBoundsRef = collect(tblBoundsRef),
  dfMetrics = dfMetrics
)

lCharts
```

## Using workflows

Unfortunately the data specfication checks from gsm.core will not work on lazy tables. We therefore need to skip the 
mapping workflow and ignore the warnings for the missing specifications.

We might also scrub some of the specifications from the yaml files to begin with.

```{r}
lMapped <- list(
  Mapped_AE = dbPortfolio$Raw_AE,
  Mapped_Visit = dbPortfolio$Raw_VISIT,
  Mapped_SUBJ = dbPortfolio$Raw_SUBJ %>%
    filter(enrollyn == "Y"),
  Mapped_StudyRef = tbl(con, "StudyRef")
)

metrics_wf <- gsm.core::MakeWorkflowList(
  strNames = "kri0001.yaml",
  strPath = system.file("workflow/2_metrics", package = "gsm.studykri"),
  strPackage = NULL
)

lAnalyzed <- gsm.core::RunWorkflows(lWorkflows = metrics_wf, lData = lMapped)

reporting_wf <- gsm.core::MakeWorkflowList(
  strNames = c("Bounds", "BoundsRef", "Input", "Join", "Metrics", "Results"),
  strPath = system.file("workflow/3_reporting", package = "gsm.studykri")
)

lReporting <- gsm.core::RunWorkflows(
  lWorkflows = reporting_wf,
  lData = c(
    lMapped,
    list(
      lAnalyzed = lAnalyzed,
      lWorkflows = metrics_wf
    )
  )
)

lReporting
```

Next we collect the data and create the charts.

```{r}
lCharts <- gsm.studykri::MakeCharts_StudyKRI(
  dfResults = collect(lReporting$Reporting_Results),
  dfBounds = collect(lReporting$Reporting_Bounds),
  dfBoundsRef = collect(lReporting$Reporting_BoundsRef),
  dfMetrics = lReporting$Reporting_Metrics
)

lCharts
```
